Hierarchically Supervised Latent Dirichlet Allocation

### 2011-05-12 paper/main.tex
Rearranged some stuff and started to NIPS-ify

\title{Supervised Topic Modeling in Clinical Text}

\author{
Perotte A\hspace{1cm} Li Y \hspace{1cm} Pivovarov R \hspace{1cm} Weiskopf N\hspace{1cm} Wood F\\
Columbia University, New York, NY 10027, USA \\
\texttt{\{\{ajp9009, yil7003, rip7002, ngw7001\}@dbmi., fwood@stat.\}columbia.edu}
%\texttt{pfau@neurotheory.columbia.edu} 
%\texttt{\{bartlett,fwood\}@stat.columbia.edu} 
}

@@

\begin{abstract}
Current medical record keeping technology relies heavily upon human
capacity to effectively summarize and infer information from free-text
physician notes. We propose a novel method to suggest diagnostic code
assignment for patient visits, based upon narrative medical notes.
We applied a supervised latent Dirichlet allocation model to a corpora
of free-text medical notes from NewYork - Presbyterian Hospital to
infer a set of specific ICD-9 codes for each patient note. Evaluation
of the predictions were conducted by comparison to a gold-standard
set of ICD-9s assigned to a set of patient notes. 
\end{abstract}

@@

\label{sec:introduction} Despite the growing emphasis on meaningful
use of technology in medicine, many aspects of medical record-keeping
remain a manual process. Diagnostic coding for billing and insurance
purposes is often handled by professional medical coders who must
explore a patient's extensive clinical record before assigning the
proper codes. So while electronic health records (EHRs) should be
adopted by most medical institutions within the next several years,
largely due to the provisions of HITECH under the American Recovery
and Reinvestment Act \citep{Blumenthal2009}, there has been little
movement forward in automating medical coding.

@@

While there have been some attempts to automatically classify groups
of patients as a potential preliminary step to ICD-9 code assignment
\citep{Ruch2008,FreitasJunior2006,RibeiroNeto2001,Brown2006}, fully
automatic assignment of ICD-9 codes to medical text became a more
prevalent research topic only in the last few years. A subset of earlier
work proposed various methods on small corpora, based on a few specific
diseases \citep{Rao2003} but the most recent and promising work on
the subject was inspired by the 2007 Medical NLP Challenge: \textquotedblleft{}International
Challenge: Classifying Clinical Free Text Using Natural Language Processing\textquotedblright{}
(website). Most of the classification strategies included word matching
and rule-based algorithms. \citep{Goldstein2007,Crammer2007,Farkas2008}.
The data set given to the participants consisted only of documents
that were 1-2 lines each and all of the documents were radiology reports
- clearly limiting the scope of potential ICD-9 codes which could
be assigned. The only paper which has attempted to work with a document
scope as large as ours was the 2008 Lita et al publication \citep{Lita2008}.
Lita proposed support vector machine and bayesian ridge regression
methods to assign appropriate labels to the documents but did not
utilize the ICD-9 hierarchy to leverage more comprehensive predictions.

@@

@@

@@

>>>>

\title{Hierarchically Supervised Latent Dirichlet Allocation}
%Supervised Topic Modeling in Clinical Text}

\author{
Adler Perotte\hspace{1cm} Nicholas Bartlett \hspace{1cm} Noemie Elhaddad \hspace{1cm} Frank Wood\\
Columbia University, New York, NY 10027, USA \\
\texttt{\{ajp9009@dbmi,bartlett@stat,noemie@dbmi,fwood@stat\}.columbia.edu}
%\texttt{pfau@neurotheory.columbia.edu} 
%\texttt{\{bartlett,fwood\}@stat.columbia.edu} 
}

@@

\begin{abstract}
The benefits of supervision in topic modeling 
%Current medical record keeping technology relies heavily upon human
%capacity to effectively summarize and infer information from free-text
%physician notes. We propose a novel method to suggest diagnostic code
%assignment for patient visits, based upon narrative medical notes.
%We applied a supervised latent Dirichlet allocation model to a corpora
%of free-text medical notes from NewYork - Presbyterian Hospital to
%infer a set of specific ICD-9 codes for each patient note. Evaluation
%of the predictions were conducted by comparison to a gold-standard
%set of ICD-9s assigned to a set of patient notes. 
\end{abstract}

@@

\label{sec:introduction}

\begin{itemize}
\item Benefits of combining human categorization information into ``topic models''
\item 
\end{itemize}

% Informatics journal paper
%
% Despite the growing emphasis on meaningful
%use of technology in medicine, many aspects of medical record-keeping
%remain a manual process. Diagnostic coding for billing and insurance
%purposes is often handled by professional medical coders who must
%explore a patient's extensive clinical record before assigning the
%proper codes. So while electronic health records (EHRs) should be
%adopted by most medical institutions within the next several years,
%largely due to the provisions of HITECH under the American Recovery
%and Reinvestment Act \citep{Blumenthal2009}, there has been little
%movement forward in automating medical coding.

@@

@@

% generalize to arbitrary codes (not just ICD-9's

@@

\section{Experiments}

@@

% move to related work?
While there have been some attempts to automatically classify groups
of patients as a potential preliminary step to ICD-9 code assignment
\citep{Ruch2008,FreitasJunior2006,RibeiroNeto2001,Brown2006}, fully
automatic assignment of ICD-9 codes to medical text became a more
prevalent research topic only in the last few years. A subset of earlier
work proposed various methods on small corpora, based on a few specific
diseases \citep{Rao2003} but the most recent and promising work on
the subject was inspired by the 2007 Medical NLP Challenge: \textquotedblleft{}International
Challenge: Classifying Clinical Free Text Using Natural Language Processing\textquotedblright{}
(website). Most of the classification strategies included word matching
and rule-based algorithms. \citep{Goldstein2007,Crammer2007,Farkas2008}.
The data set given to the participants consisted only of documents
that were 1-2 lines each and all of the documents were radiology reports
- clearly limiting the scope of potential ICD-9 codes which could
be assigned. The only paper which has attempted to work with a document
scope as large as ours was the 2008 Lita et al publication \citep{Lita2008}.
Lita proposed support vector machine and bayesian ridge regression
methods to assign appropriate labels to the documents but did not
utilize the ICD-9 hierarchy to leverage more comprehensive predictions.
